---
title: "data wrangling v2"
author: "TJ Sipin"
date: "2023-02-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(tidymodels)
```

# Data

```{r data}
data <- readRDS("~/peregrine_amazon/Restructured021623/data/annual/incidence") %>% 
  select(-c(Chikungunya, Dengue:Zika, OptTemp_Obs:Malaria_OptTemp)) %>% 
  select(Code, Year, Country, Name, has_CL, Cutaneous.Leishmaniasis, everything())
```

## Population wrangling

```{r}
colombia_population_raw <- read.csv("~/peregrine_amazon/data/colombia/raw/full_colombia_Population_raw.csv") 
brazil_population_raw <- read.csv("~/peregrine_amazon/data/brazil/raw/full_brazil_Population_raw.csv")
peru_population_raw <- read.csv("~/peregrine_amazon/data/peru/raw/full_peru_Population_raw.csv")

population_cleaner <- function(raw_data){
   data <- raw_data %>% 
     mutate(Year = str_sub(system.index, 5, 8) %>% 
              as.integer()) %>% 
     rename(Population = sum,
            Code = MuniCode) %>% 
     select(-c(system.index, .geo)) 
   data
}

colombia_population <- population_cleaner(colombia_population_raw)
brazil_population <- population_cleaner(brazil_population_raw)
peru_population <- population_cleaner(peru_population_raw)

saveRDS(colombia_population, "~/peregrine_amazon/data/colombia/processed/full_colombia_Population")
saveRDS(brazil_population, "~/peregrine_amazon/data/brazil/processed/full_brazil_Population")
saveRDS(peru_population, "~/peregrine_amazon/data/peru/processed/full_peru_Population")

```




## Presence- and absence-municipios

```{r}
monthly <- read.csv("~/peregrine_amazon/data/monthly_df.csv")

monthly_incidence <- monthly %>% 
  filter(Code %in% data$Code) %>%
  mutate(has_CL = case_when(Cutaneous.Leishmaniasis > 0 ~ 1,
                            Cutaneous.Leishmaniasis == 0 ~ 0)) %>% 
  group_by(Code, Year) %>% 
  mutate(CL_in_country_year = case_when(sum(has_CL) > 0 ~ 1,
                                        sum(has_CL) == 0 ~ 0)) %>% 
  ungroup()


monthly_incidence_v2 <- monthly_incidence %>% 
  select(Country, Year, CL_in_country_year) %>% 
  group_by(Country, Year) %>% 
  summarise(percent_CL = sum(CL_in_country_year)/n()) %>% 
  ungroup()

monthly_incidence_per_year_country <- ggplot(monthly_incidence_v2) +
  geom_line(aes(Year, percent_CL, color = Country)) + 
  ylim(c(0,1))

# Full

monthly_incidence_full <- monthly %>% 
  # remove this filter below
  # filter(Code %in% data$Code) %>% 
  mutate(has_CL = case_when(Cutaneous.Leishmaniasis > 0 ~ 1,
                            Cutaneous.Leishmaniasis == 0 ~ 0)) %>% 
  group_by(Code, Year) %>% 
  mutate(CL_in_country_year = case_when(sum(has_CL) > 0 ~ 1,
                                        sum(has_CL) == 0 ~ 0)) %>% 
  ungroup()

monthly_incidence_full_v2 <- monthly_incidence_full %>% 
  select(Country, Year, CL_in_country_year) %>% 
  group_by(Country, Year) %>% 
  summarise(percent_CL = sum(CL_in_country_year)/n()) %>% 
  ungroup()

monthly_incidence_full_per_year_country <- ggplot(monthly_incidence_full_v2) +
  geom_line(aes(Year, percent_CL, color = Country)) + 
  ylim(c(0,1))
```


## Introduce delta variables

Take delta variables of $x_{ijt}$ with respect to year $t$. Let $t_0 = 2000$: 

$$
\Delta^{t_0}_{x_{ij}} = \Delta^{2000}_{x_{ij}} = \frac{x_{ijt} - x_{ijt_0}}{t - t_0} = \frac{x_{ijt} - x_{ij2000}}{t - 2000},
$$
where $i$ denotes explanatory variable $i$ and $j$ denotes municipio $j$.


```{r}
data_v2 <- data %>% 
  group_by(Code) %>% 
  mutate(delta_2000_pland_forest = pland_forest - pland_forest[1],
         delta_2000_area_mn_forest = area_mn_forest - area_mn_forest[1],
         delta_2000_te_forest = te_forest - te_forest[1],
         delta_2000_enn_mn_forest = enn_mn_forest - enn_mn_forest[1]) %>% 
  ungroup()

data_v2 %>% 
  select(Year, pland_forest, delta_2000_pland_forest, 
         area_mn_forest, delta_2000_area_mn_forest, 
         te_forest, delta_2000_te_forest, enn_mn_forest, 
         delta_2000_enn_mn_forest, everything())
```


Now get 1- and 5-year delta variables.

```{r}
data_v3 <- data_v2 %>% 
  group_by(Code) %>% 
  mutate(delta_5_pland_forest = pland_forest - lag(pland_forest, 5),
         delta_5_area_mn_forest = area_mn_forest - lag(area_mn_forest, 5),
         delta_5_te_forest = te_forest - lag(te_forest, 5),
         delta_5_enn_mn_forest = enn_mn_forest - lag(enn_mn_forest, 5),
         delta_1_pland_forest = pland_forest - lag(pland_forest, 1),
         delta_1_area_mn_forest = area_mn_forest - lag(area_mn_forest, 1),
         delta_1_te_forest = te_forest - lag(te_forest, 1),
         delta_1_enn_mn_forest = enn_mn_forest - lag(enn_mn_forest, 1)) %>% 
  ungroup()
```

## Get superlative variables

That is, get temperatures and precipitations at the hottest, coldest, wettest, driest quarters of the year. This will require the monthly data.

```{r}
monthly_seasons <- monthly %>% 
  select(Country, Month, Year, Precip, LST_Day, LST_Night) %>% 
  pivot_longer(cols = Precip:LST_Night,
               names_to = "var")

# temperature

monthly_seasons_temp <- monthly_seasons %>% 
  filter(var %in% c("LST_Day", "LST_Night"))

ggplot(monthly_seasons_temp, 
       aes(x = Year, y = value, color = var)) + 
  geom_point() +
  geom_smooth(aes(color = Year))

```

Some limitations include varying wet and dry seasons due to El Niño and La Niña. Some assumptions to be made are to assume a constant wet and dry season.


## Incidence data

```{r}

```

