## Data and Packages

```{r}
knitr::opts_chunk$set(eval = F)
library(dplyr)
library(spatialsample)
library(sf)
library(ggplot2)
library(tidyr)


# re-run after each R session restarts
df <- readRDS("~/peregrine_amazon/data/annual/aad_geometry.rds")

df_v2 <- df %>% 
  group_by(Country, Code, Name) %>% 
  mutate(inc_temp = LST_Day > dplyr::lag(LST_Day, n=1)) %>% 
  mutate(inc_precip = Precip > dplyr::lag(Precip, n=1),
         inc_CL = Cutaneous.Leishmaniasis > dplyr::lag(Cutaneous.Leishmaniasis),
         precip_cases = case_when(inc_precip & inc_CL ~ "precip increase, CL increase",
                                  !inc_precip & inc_CL ~ "precip decrease, CL increase",
                                  !inc_precip & !inc_CL ~ "precip decrease, CL decrease",
                                  inc_precip & !inc_CL ~ "precip increase, CL decrease"),
         temp_cases   = case_when(inc_temp & inc_CL ~ "temp increase, CL increase",
                                  !inc_temp & inc_CL ~ "temp decrease, CL increase",
                                  !inc_temp & !inc_CL ~ "temp decrease, CL decrease",
                                  inc_temp & !inc_CL ~ "temp increase, CL decrease")) %>% 
  ungroup()

df_v3 <- df_v2 %>% 
  mutate(diff_CL = Cutaneous.Leishmaniasis - dplyr::lag(Cutaneous.Leishmaniasis))

colombia_2009 <- df_v3 %>% 
  group_by(Code, Country) %>% 
  mutate(deforested.lag.1.diff = pland_forest - dplyr::lag(pland_forest, n = 1),
         deforested.lag.2.diff = pland_forest - dplyr::lag(pland_forest, n = 2),
         deforested.lag.3.diff = pland_forest - dplyr::lag(pland_forest, n = 3),
         deforested.lag.4.diff = pland_forest - dplyr::lag(pland_forest, n = 4),
         deforested.lag.0.nodiff = pland_forest,
         deforested.lag.1.nodiff = dplyr::lag(pland_forest, n = 1),
         deforested.lag.2.nodiff = dplyr::lag(pland_forest, n = 2),
         deforested.lag.3.nodiff = dplyr::lag(pland_forest, n = 3),
         deforested.lag.4.nodiff = dplyr::lag(pland_forest, n = 4)) %>% 
  ungroup() %>% 
  dplyr::filter(Year == 2009,
                Country == "Colombia") %>% 
  dplyr::select(Code, Country, Year, Name, Population, Cutaneous.Leishmaniasis,
         LST_Day, LST_Night, NDVI:pland_forest, area_mn_forest, te_forest, enn_mn_forest,
         geometry:deforested.lag.4.nodiff)
```

### Add geometries to aad
```{r}
data <- read.csv("~/MacDonald-REU-Summer-22-1/models/data/aad.csv")

colombia.sh <- read_sf("~/peregrine_amazon/data/colombia/shapefiles/Colombia_Amazonian_Municipios_Projected.shp") %>% 
  select(MPIOS, geometry) %>% 
  rename(Code = MPIOS) %>% 
  mutate(Country = "Colombia")
brazil.sh <- read_sf("~/peregrine_amazon/data/brazil/shapefiles/Final_Brazil_Amazonian_Municipios_Projected.shp") %>% 
  select(codigo_ibg, geometry) %>% 
  rename(Code = codigo_ibg) %>% 
  mutate(Country = "Brazil")
peru.sh <- read_sf("~/peregrine_amazon/data/peru/shapefiles/Final_Peru_Amazonian_Municipios_Projected.shp") %>% 
  select(IDDIST, geometry) %>% 
  rename(Code = IDDIST) %>% 
  mutate(Country = "Peru")


sh <- colombia.sh %>% 
  rbind(brazil.sh) %>% 
  rbind(peru.sh) 

ggplot(sh) +
  geom_sf(aes(color = Country))
```


```{r}
data_v2 <- data %>% 
  mutate(Code = as.character(Code)) %>% 
  filter(Code %in% sh$Code)

geom_key <- data_v2 %>% 
  select(Code, Country) %>% 
  mutate(Country.1 = Country) %>% 
  unique()

sh_v2 <- data.frame(
  Code = data_v2$Code,
  Year = data_v2$Year,
  Country = data_v2$Country
) %>% 
  full_join(sh, by = c("Code", "Country")) 

ggplot(sh_v2) +
  geom_sf(aes(color = Country))
```


```{r}

sh.geometry <- st_sfc(sh_v2$geometry)



sf <- st_set_geometry(sh_v2, sh.geometry)

ggplot(sf) +
  geom_sf(aes(fill = Country))
```

```{r}
df <- data_v2 %>% 
  full_join(sf, by = c("Code", "Country", "Year"))

df <- st_set_geometry(df, sh.geometry)

ggplot(df) +
  geom_sf(aes(fill = Country))

# saveRDS(df, "~/peregrine_amazon/data/annual/aad_geometry.rds")
```



```{r}
(
  g_CL_colombia <- ggplot(
    sf %>% 
      filter(Country == "Colombia",
             Year %in% c(2007:2017)) %>% 
      mutate(has_CL = case_when(Cutaneous.Leishmaniasis == 0 ~ "no",
                                Cutaneous.Leishmaniasis > 0 ~ "yes"))) +
    geom_sf(aes(fill = Cutaneous.Leishmaniasis)) + 
    facet_wrap(facets = 'Year')
)

(
  g_CL_brazil <- ggplot(
    sf %>% 
      filter(
             !is.na(Cutaneous.Leishmaniasis),
             Year == 2011) %>% 
      mutate(has_CL = case_when(Cutaneous.Leishmaniasis == 0 ~ "no",
                                Cutaneous.Leishmaniasis > 0 ~ "yes"))) +
    geom_sf(aes(fill = Country))
)


png("~/peregrine_amazon/Restructured021623/EDA/CL_plot_geometry.png", width = 600, height = 1200)
# quartz(width = 8.5, height = 11)
ggplot(df %>% 
         filter(!is.na(Cutaneous.Leishmaniasis))) +
  geom_sf(aes(fill = Cutaneous.Leishmaniasis)) +
  scale_fill_gradient(low = "lightblue", high = "pink") + 
  facet_wrap('Year', nrow = 3, ncol = 8) +
  coord_fixed()
# quartz.save("~/peregrine_amazon/Restructured021623/EDA/CL_plot_geometry.pdf")
dev.off()
```

```{r, fig.height=36, fig.width = 12}
df <- readRDS("~/peregrine_amazon/data/annual/aad_geometry.rds")

ggplot(df %>% 
         filter(!is.na(Cutaneous.Leishmaniasis))) +
  geom_sf(aes(fill = Cutaneous.Leishmaniasis)) +
  scale_fill_gradient(low = "darkgrey", high = "blue", trans = 'log') + 
  facet_wrap('Year', shrink = F, ncol = 2)

```

```{r}
df_v2 <- df %>% 
  group_by(Country, Code, Name) %>% 
  mutate(inc_temp = LST_Day > dplyr::lag(LST_Day, n=1)) %>% 
  mutate(inc_precip = Precip > dplyr::lag(Precip, n=1),
         inc_CL = Cutaneous.Leishmaniasis > dplyr::lag(Cutaneous.Leishmaniasis),
         precip_cases = case_when(inc_precip & inc_CL ~ "precip increase, CL increase",
                                  !inc_precip & inc_CL ~ "precip decrease, CL increase",
                                  !inc_precip & !inc_CL ~ "precip decrease, CL decrease",
                                  inc_precip & !inc_CL ~ "precip increase, CL decrease"),
         temp_cases   = case_when(inc_temp & inc_CL ~ "temp increase, CL increase",
                                  !inc_temp & inc_CL ~ "temp decrease, CL increase",
                                  !inc_temp & !inc_CL ~ "temp decrease, CL decrease",
                                  inc_temp & !inc_CL ~ "temp increase, CL decrease")) %>% 
  ungroup()
```

```{r, fig.height=36, fig.width = 12}
# precip cases plot; lag=1
ggplot(df_v2) +
  geom_sf(aes(fill = (precip_cases))) +
  facet_wrap('Year', shrink = F, ncol = 2)
```

```{r, fig.height=36, fig.width = 12}
# precip cases plot; lag=2
ggplot(df_v2) +
  geom_sf(aes(fill = (precip_cases))) +
  facet_wrap('Year', shrink = F, ncol = 2)
```

```{r, fig.height=36, fig.width = 12}
# precip cases plot; lag=2
ggplot(df_v2) +
  geom_sf(aes(fill = (temp_cases))) +
  facet_wrap('Year', shrink = F, ncol = 2)
```

```{r, fig.height = 24, fig.height = 8}

ggplot(df_v2 %>% 
         dplyr::filter(temp_cases == "temp increase, CL increase")) +
  geom_sf(aes(fill = LST_Day)) + 
  facet_wrap("Year", shrink = F, ncol = 2)
```

```{r, fig.height = 24, fig.width = 18}
ggplot(df_v2,
       aes(x = LST_Day,
           y = log(Cutaneous.Leishmaniasis + 1))) +
  geom_point() +
  geom_smooth(aes(color = temp_cases)) +
  facet_wrap("Year", ncol = 2,
             scales = "free_y")
```


```{r, fig.height = 36, fig.width = 4}
ggplot(df_v2) +
  geom_boxplot(aes(x = LST_Day,
                   y = log(Cutaneous.Leishmaniasis + 1),
                   color = temp_cases)) +
  facet_wrap("Year", ncol = 1,
             scales = "fixed")
```

```{r}
df_v3 <- df_v2 %>% 
  mutate(diff_CL = Cutaneous.Leishmaniasis - dplyr::lag(Cutaneous.Leishmaniasis))

ggplot(df_v3,
       aes(x = Year,
           y = diff_CL)) +
  geom_smooth(aes(color = temp_cases))

ggplot(df_v3,
       aes(x = Year,
           y = diff_CL)) +
  geom_smooth(aes(color = precip_cases))
```

#### Paneled graphs

In this section, we panel the rows by Country $j \in \{c, b, p\}$ where $j = c$ denotes Colombia, $j = b$ denotes Brazil, and $j = p$ denotes Peru.


```{r}
df_precip <- df_v3 %>% 
  group_by(Code, Country) %>% 
  mutate(precip.lag.1.diff = Precip - dplyr::lag(Precip, n = 1),
         precip.lag.2.diff = Precip - dplyr::lag(Precip, n = 2),
         precip.lag.3.diff = Precip - dplyr::lag(Precip, n = 3),
         precip.lag.4.diff = Precip - dplyr::lag(Precip, n = 4),
         precip.lag.0.nodiff = Precip,
         precip.lag.1.nodiff = dplyr::lag(Precip, n = 1),
         precip.lag.2.nodiff = dplyr::lag(Precip, n = 2),
         precip.lag.3.nodiff = dplyr::lag(Precip, n = 3),
         precip.lag.4.nodiff = dplyr::lag(Precip, n = 4)) %>% 
  ungroup() %>% 
  select(Code, Year, Country, Cutaneous.Leishmaniasis, contains("precip.lag")) %>% 
  pivot_longer(cols = c(contains(".diff"), contains(".nodiff"))) %>% 
  mutate(nodiff = grepl(".nodiff", name))
```

```{r, fig.height = 32, fig.width = 12}
ggplot(df_precip %>% 
         select(Code, Country, Year, Cutaneous.Leishmaniasis, name, nodiff, value) %>% 
         dplyr::filter(nodiff),
       aes(x = value,
           y = Cutaneous.Leishmaniasis)) +
  geom_smooth(aes(color = name)) +
  facet_grid(cols = vars(Country),
             rows = vars(Year), 
             scales = "fixed")
```

There is a small hump around precipitation values $precip_{jt} = 2000$ in years $t = 2000,\dots, 2009$, and again from years $t = 2015, \dots, 2018$.


```{r}
df_temp <- df_v3 %>% 
  group_by(Code, Country) %>% 
  mutate(temp.lag.1.diff = LST_Day - dplyr::lag(LST_Day, n = 1),
         temp.lag.2.diff = LST_Day - dplyr::lag(LST_Day, n = 2),
         temp.lag.3.diff = LST_Day - dplyr::lag(LST_Day, n = 3),
         temp.lag.4.diff = LST_Day - dplyr::lag(LST_Day, n = 4),
         temp.lag.0.nodiff = LST_Day,
         temp.lag.1.nodiff = dplyr::lag(LST_Day, n = 1),
         temp.lag.2.nodiff = dplyr::lag(LST_Day, n = 2),
         temp.lag.3.nodiff = dplyr::lag(LST_Day, n = 3),
         temp.lag.4.nodiff = dplyr::lag(LST_Day, n = 4)) %>% 
  ungroup() %>% 
  select(Code, Year, Country, Cutaneous.Leishmaniasis, contains("temp.lag")) %>% 
  pivot_longer(cols = c(contains(".diff"), contains(".nodiff"))) %>% 
  mutate(nodiff = grepl(".nodiff", name))
```


```{r, fig.height = 32, fig.width = 18}
ggplot(df_temp %>% 
         select(Code, Year, Country, Cutaneous.Leishmaniasis, name, nodiff, value) %>% 
         dplyr::filter(nodiff),
       aes(x = value,
           y = Cutaneous.Leishmaniasis)) +
  geom_smooth(aes(color = name)) +
  facet_grid(cols = vars(Country),
             rows = vars(Year), 
             scales = "fixed")
```

Note:

- The high variation in 2009 Colombia
- The characteristic higher incidence in the lower tail for Brazil at $LST_{bt} = 25$ for all $t$.


```{r}
population_df <- df_v3 %>% 
  group_by(Country, Year) %>% 
  mutate(population_percentile = percent_rank(Population),
         CL_percentile = percent_rank(Cutaneous.Leishmaniasis)) %>% 
  ungroup() %>% 
  select(Code, Country, Year, population_percentile, CL_percentile, Population, Cutaneous.Leishmaniasis, precip_cases)
```

```{r, fig.height = 12, fig.width = 12}
# percentile plot
ggplot(population_df, 
       aes(x = population_percentile,
           y = CL_percentile)) +
  geom_smooth(aes(color = Country)) +
  # facet_grid(rows = vars(Year)) +
  facet_wrap(~ Year) +
  coord_fixed()
```

```{r, fig.height = 12, fig.width = 12}
# regular plot
ggplot(population_df, 
       aes(x = Population,
           y = Cutaneous.Leishmaniasis)) +
  geom_smooth(aes(color = Country)) +
  # facet_grid(rows = vars(Year)) +
  facet_wrap(~ Year) 
```

The percentile rank plots give more clear information, but perhaps I can configure a way to make the raw values plots easier to read.

```{r, fig.height = 12, fig.width = 12}
ggplot(population_df, 
       aes(x = Population,
           y = log(Cutaneous.Leishmaniasis + 1))) +
  geom_smooth(aes(color = Country)) +
  # facet_grid(rows = vars(Year)) +
  facet_wrap(~ Year) 

ggplot(population_df, 
       aes(x = Population,
           y = log(Cutaneous.Leishmaniasis + 1))) +
  geom_smooth(aes(color = Country)) +
  # facet_grid(rows = vars(Year)) +
  facet_wrap(~ Year) 
```
### Deforestation plots

```{r}
df_deforested <- df_v3 %>% 
  group_by(Code, Country) %>% 
  mutate(deforested.lag.1.diff = pland_forest - dplyr::lag(pland_forest, n = 1),
         deforested.lag.2.diff = pland_forest - dplyr::lag(pland_forest, n = 2),
         deforested.lag.3.diff = pland_forest - dplyr::lag(pland_forest, n = 3),
         deforested.lag.4.diff = pland_forest - dplyr::lag(pland_forest, n = 4),
         deforested.lag.0.nodiff = pland_forest,
         deforested.lag.1.nodiff = dplyr::lag(pland_forest, n = 1),
         deforested.lag.2.nodiff = dplyr::lag(pland_forest, n = 2),
         deforested.lag.3.nodiff = dplyr::lag(pland_forest, n = 3),
         deforested.lag.4.nodiff = dplyr::lag(pland_forest, n = 4)) %>% 
  ungroup() %>% 
  select(Code, Year, Country, Cutaneous.Leishmaniasis, contains("deforested.lag")) %>% 
  pivot_longer(cols = c(contains(".diff"), contains(".nodiff"))) %>% 
  mutate(nodiff = grepl(".nodiff", name))
```

```{r, fig.height = 32, fig.width = 18}
ggplot(df_deforested %>% 
         dplyr::filter(nodiff == T),
       aes(x = value,
           y = Cutaneous.Leishmaniasis)) +
  geom_smooth(aes(color = name)) + 
  facet_grid(cols = vars(Country),
             rows = vars(Year),
             scales = "free_x")
```


### Time series 

```{r}
library(forecast)
```   

## 2009

There's a common theme that CL incidence in 2009, particularly in Colombia, has high variance. This can be investigated further.

```{r}
colombia_2009 <- df_v3 %>% 
  group_by(Code, Country) %>% 
  mutate(deforested.lag.1.diff = pland_forest - dplyr::lag(pland_forest, n = 1),
         deforested.lag.2.diff = pland_forest - dplyr::lag(pland_forest, n = 2),
         deforested.lag.3.diff = pland_forest - dplyr::lag(pland_forest, n = 3),
         deforested.lag.4.diff = pland_forest - dplyr::lag(pland_forest, n = 4),
         deforested.lag.0.nodiff = pland_forest,
         deforested.lag.1.nodiff = dplyr::lag(pland_forest, n = 1),
         deforested.lag.2.nodiff = dplyr::lag(pland_forest, n = 2),
         deforested.lag.3.nodiff = dplyr::lag(pland_forest, n = 3),
         deforested.lag.4.nodiff = dplyr::lag(pland_forest, n = 4)) %>% 
  ungroup() %>% 
  dplyr::filter(Year == 2009,
                Country == "Colombia") %>% 
  dplyr::select(Code, Country, Year, Name, Population, Cutaneous.Leishmaniasis,
         LST_Day, LST_Night, NDVI:pland_forest, area_mn_forest, te_forest, enn_mn_forest,
         geometry:deforested.lag.4.nodiff)

summary(colombia_2009)
```

This year was particularly hotter and drier than 2008.

```{r}
library(rgdal)
library(raster)
library(ggplot2)
library(rgeos)
library(mapview)
library(leaflet)
library(broom) # if you plot with ggplot and need to turn sp data into dataframes

# for loading our data
library(sf)
# for plotting
library(lattice)
library(leafpop)
library(mapview)
# library(vapoRwave)
library(viridis)
```

```{r}
mapview(colombia_2009,
        zcol = "diff_CL")
```

#### pop-up graph

```{r}
# source: https://bookdown.org/nicohahn/making_maps_with_r5/docs/mapview.html#using-mapview-to-create-maps
p_temp <- xyplot(diff_CL ~ LST_Day, data = colombia_2009, col = "grey", pch = 20, cex = 2)
p_temp <- mget(rep("p_temp", nrow(colombia_2009)))

clr <- rep("grey", nrow(colombia_2009))
alp <- rep(0.2, nrow(colombia_2009))
p_temp <- lapply(1:length(p_temp), function(i) {
  clr[i] <- "red"
  alp[i] <- 1
  update(p_temp[[i]], col = clr, alpha = alp)
})

mapview(colombia_2009,
        zcol = "diff_CL", 
        popup = popupGraph(p_temp))
```

```{r}
colombia_2009_flag <- colombia_2009 %>% 
  mutate(flag = diff_CL > 9)

# colombia_2009_flag_longer <- colombia_2009_flag %>% 
#   pivot_longer(c(deforested.lag.1.diff:deforested.lag.4.nodiff))

p_forest.lag.1.diff <- xyplot(diff_CL ~ deforested.lag.1.diff, data = colombia_2009_flag, col = "grey", pch = 20, cex = 2)
p_forest.lag.1.diff <- mget(rep("p_forest.lag.1.diff", nrow(colombia_2009_flag)))

clr <- rep("grey", nrow(colombia_2009_flag))
alp <- rep(0.2, nrow(colombia_2009_flag))
p_forest.lag.1.diff <- lapply(1:length(p_forest.lag.1.diff), function(i) {
  clr[i] <- "red"
  alp[i] <- 1
  update(p_forest.lag.1.diff[[i]], col = clr, alpha = alp)
})

colombia_2009_flag.lag.1.diff <- mapview(colombia_2009_flag, zcol = 'diff_CL', popup = popupGraph(p_forest.lag.1.diff))
# saveRDS(colombia_2009_flag.lag.1.diff, "~/peregrine_amazon/Restructured021623/EDA/plots/colombia_2009_flag.lag.1.diff")
```

```{r}
p_forest.lag.2.diff <- xyplot(diff_CL ~ deforested.lag.2.diff, data = colombia_2009_flag, col = "grey", pch = 20, cex = 2)
p_forest.lag.2.diff <- mget(rep("p_forest.lag.2.diff", nrow(colombia_2009_flag)))

clr <- rep("grey", nrow(colombia_2009_flag))
alp <- rep(0.2, nrow(colombia_2009_flag))
p_forest.lag.2.diff <- lapply(1:length(p_forest.lag.2.diff), function(i) {
  clr[i] <- "red"
  alp[i] <- 1
  update(p_forest.lag.2.diff[[i]], col = clr, alpha = alp)
})

colombia_2009_flag.lag.2.diff <- mapview(colombia_2009_flag, zcol = 'diff_CL', popup = popupGraph(p_forest.lag.2.diff))
# saveRDS(colombia_2009_flag.lag.2.diff, "~/peregrine_amazon/Restructured021623/EDA/plots/colombia_2009_flag.lag.2.diff")
```


High variance likely caused by the one outlier La Macarena (50350). It seems that the surrounding neighborhood of La Macarena also had an increase in CL incidence. Let's try to figure out what's causing that. We can flag the municipios that have a CL incidence increase of more than 9.

```{r}
p_forest.lag.0.nodiff <- xyplot(diff_CL ~ deforested.lag.0.nodiff, data = colombia_2009_flag, col = "grey", pch = 20, cex = 2)
p_forest.lag.0.nodiff <- mget(rep("p_forest.lag.0.nodiff", nrow(colombia_2009_flag)))

clr <- rep("grey", nrow(colombia_2009_flag))
alp <- rep(0.2, nrow(colombia_2009_flag))
p_forest.lag.0.nodiff <- lapply(1:length(p_forest.lag.0.nodiff), function(i) {
  clr[i] <- "red"
  alp[i] <- 1
  update(p_forest.lag.0.nodiff[[i]], col = clr, alpha = alp)
})

colombia_2009_flag.lag.0.nodiff <- mapview(colombia_2009_flag, zcol = 'diff_CL', popup = popupGraph(p_forest.lag.0.nodiff))
# saveRDS(colombia_2009_flag.lag.0.nodiff, "~/peregrine_amazon/Restructured021623/EDA/plots/colombia_2009_flag.lag.0.nodiff")
```


```{r, eval=T}
readRDS("~/peregrine_amazon/Restructured021623/EDA/plots/colombia_2009_flag.lag.0.nodiff")
readRDS("~/peregrine_amazon/Restructured021623/EDA/plots/colombia_2009_flag.lag.1.diff")
readRDS("~/peregrine_amazon/Restructured021623/EDA/plots/colombia_2009_flag.lag.2.diff")
```


